Simulation of 2D physics of objects captured by web camera using OpenCV and Box2D

### 2011-02-08 latex/paper.tex
dopisany obsah, okrem futere work a abstractu

\documentclass{ifacconf}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{natbib}

@@

\begin{document}
\begin{frontmatter}
\title{Simulation of 2D physics of hand drawn objects using OpenCV and Box2D}
\author[Bratislava]{Michal Sedlák}
\address[Bratislava]{Faculty of Electrical Engineering and Information
Technology, Slovak University of Technology, Ilkovičova 3, 812 19 Bratislava,
Slovakia
\\
(e-mail: michal.sedlak@stuba.sk)}
\begin{abstract}
%TODO Napisat abstrakt

@@

Simple image capture is shown in Listing \ref{Capture}.
\begin{lstlisting}[caption=Query image frame from web camera, label=Capture]
self.camera = cv.CaptureFromCAM(-1) %\label{Capture:init})
self.image = cv.QueryFrame(self.camera) %\label{Capture:query})
self.DetectOutline(self.image)
\end{lstlisting}
In line \ref{Capture:init} of listing \ref{Capture} we initialize our web
camera. In variable camera is allocated and initialized object that can query
camera for new image. Then as we see in \ref{Capture:query} we can get the
image from camera and store it in the variable named image. Now when we have
image data stored in the variable, we can process data to find outlines.
\begin{lstlisting}[caption=Outline detection, label=Detect]
def DetectOutline(self, image):
  image_size = cv.GetSize(image)
  grayscale = cv.CreateImage(image_size, 8, 1)	%\label{Detect:grayscale})
  cv.CvtColor(image, grayscale, cv.CV_BGR2GRAY)
  cv.EqualizeHist(grayscale, grayscale)	%\label{Detect:equalize})
  storage = cv.CreateMemStorage(0)
  cv.Threshold(grayscale, grayscale, 50, 255, cv.CV_THRESH_BINARY) %\label{Detect:threshold})
  self.contours = cv.FindContours(grayscale, %\label{Detect:contours})
    cv.CreateMemStorage(),
    cv.CV_RETR_TREE,
    cv.CV_CHAIN_APPROX_SIMPLE)
  if len(self.contours) > 0:
    self.contours = cv.ApproxPoly (self.contours, %\label{Detect:approx})
      storage,
      cv.CV_POLY_APPROX_DP,
      1.5,
      1)
  return self.contours
\end{lstlisting}
In function in Listing \ref{Detect} is shown how to find outlines of objects in
image. We convert image to gray scale as seen on line \ref{Detect:grayscale}.
Then we run histogram equalization (line:~\ref{Detect:equalize}).
Equalization makes objects better visible and gives better output for
thresholding (line:~\ref{Detect:threshold}) which makes black and white image
prepared for outline detection (line:~\ref{Detect:contours}).
%TODO Doplnit obrazok s normalizaciou histogramu
%TODO Doplnit obrazok s thresholdom
After outline detection we have tree of contours stored in the variable
self.contours. These trees are iterable objects sorted from outer to inner
outline connected by property h\_next and v\_next that we will describe in
paragraph about creation of objects from outlines.

Contour can be very complicated and consist of thousands of points, which could
cause too complicated objects. It is time demanding to simulate complicated
objects, that is why we use polynomial approximation of the contour points.
(line:~\ref{Detect:approx}).

Now we have all outlines stored in the outline tree structure, so we can create
objects and apply a physics.
%TODO Doplnit obrazok so zvyraznenymi outlineami

@@

Listing of the function CreateObject() continous in Listing \ref{Objects2}. This
part of function creates dynamic objects inside the outer contour. In this part
we prepare list for bodies of objects, so we can modify objects that are already
created or objects that we want append new shapes.
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects2,caption=Creation of objects] if v == 1:
    try:
      body = self.objectBodies[h]
    except:
      body = self.world.CreateBody(bd)
      self.objectBodies[h] = body
\end{lstlisting}
\subsection{Tesselation}
Box2D supports only collisions between convex objects and contours of objects
captured by camera are not convex we have to break outlines to convex polygons.
There is more ways how to break concave objects. We chose the 2D constrained
Delaunay triangulation algorithm implemented by poly2tri Python
library[\cite{Delauanay}]. Function CreateObject() continous in Listing~
\ref{Objects3}
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects3,caption=Creation
 of objects]
    polyline = []
    for (x,y) in cont:
      polyline.append(p2t.Point(x,y)) %\label{Objects:polyline})
    cdt = p2t.CDT(polyline) %\label{Objects:p2t})
    triangles = cdt.triangulate()
    for t in triangles:
      x1 = t.a.x/30.0
      y1 = t.a.y/30.0
      x2 = t.b.x/30.0
      y2 = t.b.y/30.0
      x3 = t.c.x/30.0
      y3 = t.c.y/30.0
      if math.hypot(x2-x1,y2-y1)<0.1: %\label{Objects:minSize1})
        x2 = x2 + math.copysign(0.1, x2-x1)
        y2 = y2 + math.copysign(0.1, y2-y1)
      if math.hypot(x3-x2,y3-y2)<0.1:
        x3 = x3 + math.copysign(0.1, x3-x2)
        y3 = y3 + math.copysign(0.1, y3-y2)
      if math.hypot(x1-x3,y1-y3)<0.1:
        x1 = x1 + math.copysign(0.1, x1-x3)
        y1 = y1 + math.copysign(0.1, y1-y3) %\label{Objects:minSize2})
      poly=box2d.b2PolygonDef()  %\label{Objects:polygondef})
      poly.setVertices(((x1, y1), (x2, y2), (x3, y3)))
      poly.density = 1.0
      poly.restitution = 0.0
      poly.friction = 0.0
      body.CreateShape(poly)  %\label{Objects:createshape})
      body.SetMassFromShapes() %\label{Objects:setMass})
\end{lstlisting}
Objects:polyline
Objects:p2t
Objects:minSize1
Objects:minSize2
Objects:polygondef
Objects:createshape
Objects:setMass

\subsection{Implementation}
%TODO dat sem zdrojovy kod a popisat ho
\section{Future work}
Identifikacia objektov
Sledovanie objektov a morfing 
Interakcia hybucich sa objektov zachytenych kamerov s Box2D reprezentaciou

>>>>

\documentclass{ifacconf}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{graphicx}
\usepackage{natbib}

@@

\begin{document}
\begin{frontmatter}
\title{Simulation of 2D physics of objects captured by web camera using OpenCV
and Box2D}
\author[Bratislava]{Michal Sedlák}
\address[Bratislava]{Faculty of Electrical Engineering and Information
Technology, Slovak University of Technology, Ilkovičova 3, 812 19 Bratislava,
Slovakia
\\
(e-mail: michal.sedlak@stuba.sk)}
\begin{abstract}
%TODO Napisat abstrakt

@@

Simple image capture is shown in Listing \ref{Capture}.
\begin{lstlisting}[caption=Query image frame from web camera, label=Capture]
self.camera = cv.CaptureFromCAM(-1) %\label{Capture:init})
self.image = cv.QueryFrame(self.camera) %\label{Capture:query})
self.DetectOutline(self.image)
\end{lstlisting}
In line \ref{Capture:init} of listing \ref{Capture} we initialize our web
camera. In variable camera is allocated and initialized object that can query
camera for new image. Then as we see in \ref{Capture:query} we can get the
image from camera and store it in the variable named image. Capured image
is shown in Figure \ref{fig:captured}.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/1test.png}
\caption{Image captured form camera}
\label{fig:captured}
\end{figure}

Now when we have
image data stored in the variable, we can process data to find outlines.
\begin{lstlisting}[caption=Outline detection, label=Detect]
def DetectOutline(self, image):
  image_size = cv.GetSize(image)
  grayscale = cv.CreateImage(image_size, 8, 1)	%\label{Detect:grayscale})
  cv.CvtColor(image, grayscale, cv.CV_BGR2GRAY)
  cv.EqualizeHist(grayscale, grayscale)	%\label{Detect:equalize})
  storage = cv.CreateMemStorage(0)
  cv.Threshold(grayscale, grayscale, 50, 255, cv.CV_THRESH_BINARY) %\label{Detect:threshold})
  self.contours = cv.FindContours(grayscale, %\label{Detect:contours})
    cv.CreateMemStorage(),
    cv.CV_RETR_TREE,
    cv.CV_CHAIN_APPROX_SIMPLE)
  if len(self.contours) > 0:
    self.contours = cv.ApproxPoly (self.contours, %\label{Detect:approx})
      storage,
      cv.CV_POLY_APPROX_DP,
      1.5,
      1)
  return self.contours
\end{lstlisting}
In function in Listing \ref{Detect} is shown how to find outlines of objects in
image. We convert image to gray scale as seen on line \ref{Detect:grayscale}.

Then we run histogram equalization (line:~\ref{Detect:equalize}). Equalization
makes objects better visible and gives better output for thresholding
(line:~\ref{Detect:threshold}) which makes black and with image prepared for outline detection (line:~\ref{Detect:contours}). Output
of thresholding is shown in (Figure \ref{fig:threshold}).

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/3test-equalize-histogram.png}
\caption{Image after histogram equalization}
\label{fig:histogram}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/4test-threshold.png}
\caption{Image after thresholding}
\label{fig:threshold}
\end{figure}

After outline detection we have tree of contours stored in the variable
self.contours. These trees are iterable objects sorted from outer to inner
outline connected by property h\_next and v\_next that we will describe in
paragraph about creation of objects from outlines.

Contour can be very complicated and consist of thousands of points, which could
cause too complicated objects. It is time demanding to simulate complicated
objects, that is why we use polynomial approximation of the contour points.
(line:~\ref{Detect:approx}). Visualusation of outlines is shown in Figure:
\ref{fig:contours}.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/5test-contours.png}
\caption{Visualisation of contours}
\label{fig:contours}
\end{figure}

Now we have all outlines stored in the outline tree structure, so we can create
objects and apply a physics.

@@

Listing of the function CreateObject() continous in Listing \ref{Objects2}. This
part of function creates dynamic objects inside the outer contour. In this part
we prepare list for bodies of objects, so we can modify objects that are already
created or objects that we want append new shapes.
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects2,caption=Creation of objects] if v == 1:
    try:
      body = self.objectBodies[h]
    except:
      body = self.world.CreateBody(bd)
      self.objectBodies[h] = body
\end{lstlisting}
\subsection{Tesselation}
Box2D supports only collisions between convex objects and contours of objects
captured by camera are not convex we have to break outlines to convex polygons.
There is more ways how to break concave objects. We chose the 2D constrained
Delaunay triangulation algorithm implemented by poly2tri Python
library[\cite{Delauanay}]. Function CreateObject() continous in Listing~
\ref{Objects3}
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects3,caption=Creation
 of objects]
    polyline = []
    for (x,y) in cont:
      polyline.append(p2t.Point(x,y)) %\label{Objects:polyline})
    cdt = p2t.CDT(polyline) %\label{Objects:p2t})
    triangles = cdt.triangulate()
    for t in triangles:  %\label{Objects:scale1})
      x1 = t.a.x/30.0
      y1 = t.a.y/30.0
      x2 = t.b.x/30.0
      y2 = t.b.y/30.0
      x3 = t.c.x/30.0
      y3 = t.c.y/30.0  %\label{Objects:scale2})
      if math.hypot(x2-x1,y2-y1)<0.1: %\label{Objects:minSize1})
        x2 = x2 + math.copysign(0.1, x2-x1)
        y2 = y2 + math.copysign(0.1, y2-y1)
      if math.hypot(x3-x2,y3-y2)<0.1:
        x3 = x3 + math.copysign(0.1, x3-x2)
        y3 = y3 + math.copysign(0.1, y3-y2)
      if math.hypot(x1-x3,y1-y3)<0.1:
        x1 = x1 + math.copysign(0.1, x1-x3)
        y1 = y1 + math.copysign(0.1, y1-y3) %\label{Objects:minSize2})
      poly=box2d.b2PolygonDef()  %\label{Objects:polygondef})
      poly.setVertices(((x1, y1), (x2, y2), (x3, y3)))
      poly.density = 1.0
      poly.restitution = 0.0
      poly.friction = 0.0
      body.CreateShape(poly)  %\label{Objects:createshape})
      body.SetMassFromShapes() %\label{Objects:setMass})
\end{lstlisting}
The creation of objects continues with tesselation. We need to assigne vertices
to structure that could be understood by poly2tri library
(line:~\ref{Objects:polyline}) and we initialize the CDT object
(line:~\ref{Objects:p2t}). In next line we call function that will create
triangles from the verteces assigned before. These triangles are in image pixel
coordinates, so we need to scale them at first
(lines:~\ref{Objects:scale1}-\ref{Objects:scale2}). Now when we have triangles
scaled we need to scale the triangles that are too small to triangles with size
at least 0.1m because of speed optimalization, this is done in
lines:~\ref{Objects:minSize1}-\ref{Objects:minSize2}). Now we have set of
tirangular shapes that could be attached to body
(line:~\ref{Objects:createshape:}). Because these objects are compound objects,
we need to set the mass center and ammount to this body. We can let Box2D set
this properties based on shape information with body function
SetMassFromShapes() (line:~\ref{Objects:setMass}). Visualisation of objects is
in Figure ref{fig:objects}

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/6dynamic-objets.png}
\caption{Visualisation of objects after triangulation}
\label{fig:objects}
\end{figure}

After this we can add other objects and start simulation by function Step().
After few second of simulation are all objects on the bottom of the screen like
in Figure \ref{fig:simulation}.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{images/7dynamic-objets-simulation.png}
\caption{Visualisation of simulation}
\label{fig:simulation}
\end{figure}

\section{Future work}
Identifikacia objektov
Sledovanie objektov a morfing 
Interakcia hybucich sa objektov zachytenych kamerov s Box2D reprezentaciou
